---
title: "Region Breakdown Pt 2 - Model Baseline Switch"
output:
  html_document:
    df_print: paged
  word_document:
    reference_docx: template/template.docx
date: "Oct 31, 2023"
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, fig.width = 14, fig.height = 8)

library(tidyverse)
library(flextable)
library(janitor)
library(scales)

data <- read_csv(here::here("data/io.csv"))
data <- data %>% filter(districtName %in% c("Texas North", "Texas West", "Texas South", "Texas East","Louisiana East", "Louisiana West", "New Mexico", "Florida North", "Florida Middle", "Florida South"))
#add koon period?

postprotect <- data %>% 
  filter(postprotect) #06-24-04

postbooker <- data %>% 
  filter(postbooker) #dates are not right #12-10-2007

postgall <- data %>% 
  filter(postgall)

postreport <- data %>% 
  filter(postreport)

present <- data %>% 
  filter(present)


source(here::here("utils.R"))

```


This document attempts to replicate the Commission's "refined model" for estimating demographic disparities in sentencing. For more information about the model and for the reference regressions, see "Demographic Disparities Baseline Switch Reference Docs." For a write-up of findings, see the latest version of the Data Learnings Memo.

## White Male Baseline Regressions

```{r}
model2 <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 8) +
               agedummy +
               educ +
               citizen,
             df)
}
```

### Post-Protect

```{r}
model2_factor_protect <- model2(postprotect)

as_flextable(model2_factor_protect)
```
<br>
<br>

### Post-Booker

```{r}
model2_factor_booker <- model2(postbooker)

as_flextable(model2_factor_booker)
```

```{r, include = F}
summary(model2_factor_booker)
```
<br>
<br>

### Post-Gall

```{r}
model2_factor_gall <- model2(postgall)

as_flextable(model2_factor_gall)
```

```{r, include = F}
summary(model2_factor_gall)
```
<br>
<br>

### Post-Report

```{r}
model2_factor_report <- model2(postreport)

as_flextable(model2_factor_report)
```
<br>
<br>

### Present

```{r}
model2_factor_present <- model2(present)

as_flextable(model2_factor_present)
```

<br>
<br>

```{r, include = F}
### Coefficient Interpretations for All Periods

protect_interp <- extract_coeffs_tbl(model2_factor_protect) %>% select(var, Protect = coeffinterp)
booker_interp <- extract_coeffs_tbl(model2_factor_booker) %>% select(var, Booker = coeffinterp)
gall_interp <- extract_coeffs_tbl(model2_factor_gall) %>% select(var, Gall = coeffinterp)
report_interp <- extract_coeffs_tbl(model2_factor_report) %>% select(var, Report = coeffinterp)
present_interp <- extract_coeffs_tbl(model2_factor_present) %>% select(var, Present = coeffinterp)

interp_table <- protect_interp %>% 
  full_join(booker_interp) %>% 
  full_join(gall_interp) %>% 
  full_join(report_interp) %>% 
  full_join(present_interp) %>% 
  mutate(" " = var %>% 
           str_remove_all("TRUE") %>% 
           str_remove_all("2") %>% 
           str_remove_all("relevel\\(factor\\(racesex\\)\\, ref = 8\\)")) %>% 
  filter(!str_detect(var, "ntercept")) %>% 
select(" ", everything(), -var)



graph_table_a <- interp_table %>% 
  pivot_longer(cols = c(Protect, Booker, Gall, Report, Present), names_to = "period", values_to = "coeffinterp") %>% 
  mutate(period = case_when(
    period=="Protect" ~ 1,
    period=="Booker" ~ 2,
    period=="Gall" ~ 3,
    period=="Report" ~ 4,
    period=="Present" ~ 5,
  )) %>% 
  select(var = " ", period, coeffinterp)

list_vars <- graph_table_a %>% pull(var) %>% unique()

line_graph <- function(varname) {
  graph_table_a %>% 
    filter(var==varname) %>% 
    ggplot() +
    geom_path(aes(x = period, y = abs(coeffinterp))) +
    theme_void()
}



ft <- interp_table %>%
  mutate(Trend = map(list_vars, line_graph)) %>%
  flextable() %>%
  compose(j = c("Trend"),
          value = as_paragraph(gg_chunk(value = ., height = .25, width = .6)),
          use_dot = TRUE) %>%
  set_table_properties(width = .65, layout = "autofit") %>%
  add_footer_lines("Trend lines are not-to-scale representions of coefficient magnitude over time.")

save_as_image(ft, here::here("coefficientinterps.png"), zoom = 3, expand = 10, webshot = "webshot")



```

<br>
  
<br>
  
<br>
  
<br>
  
<br>
  
<br>
  
<br>
  
<br>
  
<br>
  
<br>
  
<br>
  
<br>
  
\newpage 
 
 
## Demographic Disparities in Sentencing: Bar Graphs

```{r}

demogbar <- function(df){
  df %>% 
    ggplot(aes(x = period, y = value, fill = period))+
    geom_bar(stat="identity")+
    geom_text(aes(label=value), vjust=0.3, size = 8.5)+ 
    labs(y = "Percent", x = "Period",
         caption = "Values not statistically significant at the 0.01 level are omitted.")+
    theme_minimal() + 
    facet_wrap(~demog)+
    scale_fill_brewer(type = "qual", palette = 6) +
    theme(strip.text = element_text(size = 25),
          legend.text = element_text(size = 20),
          legend.title = element_text(size = 22),
          axis.title.y = element_text(size = 22),
          axis.title.x = element_text(size = 22),
          plot.caption = element_text(size = 20),
          axis.text.y = element_blank(),
          axis.text.x=element_blank())
  
}

```


### White Male Baseline

```{r}

extract_coeffs_demog <- function(df){
  df %>%
    extract_coeffs() %>% 
    filter(str_detect(var, "male")) %>% 
    mutate(value = logtrans(coeff),
           demog = case_when(
            str_detect(var, "whitemale") ~ "White Male",
            str_detect(var, "whitefemale") ~ "White Female",
            str_detect(var, "blackmale") ~ "Black Male",
            str_detect(var, "blackfemale") ~ "Black Female",
            str_detect(var, "hispmale") ~ "Hispanic Male",
            str_detect(var, "hispfemale") ~ "Hispanic Female",
            str_detect(var, "othermale") ~ "Other Male",
            str_detect(var, "otherfemale") ~ "Other Female")) %>% 
    select(demog, value)
}

protect_wm <- extract_coeffs_demog(model2_factor_protect) %>% mutate(period = "Protect")
booker_wm <- extract_coeffs_demog(model2_factor_booker) %>% mutate(period = "Booker")
gall_wm <- extract_coeffs_demog(model2_factor_gall) %>% mutate(period = "Gall")
report_wm <- extract_coeffs_demog(model2_factor_report) %>% mutate(period = "Post-Report")
present_wm <- extract_coeffs_demog(model2_factor_present) %>% mutate(period = "Present")


bar_data_wm <- bind_rows(protect_wm,
                         booker_wm,
                         gall_wm,
                         report_wm,
                         present_wm) %>% 
  mutate(period = fct_relevel(period, "Protect", "Booker", "Gall", "Post-Report", "Present"))
```

```{r}
bar_data_wm %>% filter(str_detect(demog, "Female")) %>% demogbar()
```
  
  
```{r}
bar_data_wm %>% filter(str_detect(demog, "Male")) %>% demogbar()

#add in statistical significance dots
#fix position of bar labels

```
<br>
<br>

### White Female Baseline

Note the colors are scrambled slightly where no statistically significant difference in groups exists during the Protect period.

```{r}
model2_wf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 7) +
               agedummy +
               educ +
               citizen,
             df)
}
```

```{r}
model2_protect_wf <- model2_wf(postprotect)
model2_booker_wf <- model2_wf(postbooker)
model2_gall_wf <- model2_wf(postgall)
model2_report_wf <- model2_wf(postreport)
model2_present_wf <- model2_wf(present)


```

```{r}
protect_wf <- extract_coeffs_demog(model2_protect_wf) %>% mutate(period = "Protect")
booker_wf <- extract_coeffs_demog(model2_booker_wf) %>% mutate(period = "Booker")
gall_wf <- extract_coeffs_demog(model2_gall_wf) %>% mutate(period = "Gall")
report_wf <- extract_coeffs_demog(model2_report_wf) %>% mutate(period = "Post-Report")
present_wf <- extract_coeffs_demog(model2_present_wf) %>% mutate(period = "Present")


bar_data_wf <- bind_rows(protect_wf,
                         booker_wf,
                         gall_wf,
                         report_wf,
                         present_wf) %>% 
  mutate(period = fct_relevel(period, "Protect", "Booker", "Gall", "Post-Report", "Present"))

bar_data_wf %>% filter(str_detect(demog, "Female")) %>% demogbar() 
  
```
  
  
```{r}
bar_data_wf %>% filter(str_detect(demog, "Male")) %>% demogbar()
```
<br>
<br>

### Hispanic Female Baseline

```{r}
model2_hf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 3) +
               agedummy +
               educ +
               citizen,
             df)
}
```

```{r}
model2_protect_hf <- model2_hf(postprotect)
model2_booker_hf <- model2_hf(postbooker)
model2_gall_hf <- model2_hf(postgall)
model2_report_hf <- model2_hf(postreport)
model2_present_hf <- model2_hf(present)

```

```{r}
protect_hf <- extract_coeffs_demog(model2_protect_hf) %>% mutate(period = "Protect")
booker_hf <- extract_coeffs_demog(model2_booker_hf) %>% mutate(period = "Booker")
gall_hf <- extract_coeffs_demog(model2_gall_hf) %>% mutate(period = "Gall")
report_hf <- extract_coeffs_demog(model2_report_hf) %>% mutate(period = "Post-Report")
present_hf <- extract_coeffs_demog(model2_present_hf) %>% mutate(period = "Present")


bar_data_hf <- bind_rows(protect_hf,
                         booker_hf,
                         gall_hf,
                         report_hf,
                         present_hf) %>% 
  mutate(period = fct_relevel(period, "Protect", "Booker", "Gall", "Post-Report", "Present"))

bar_data_hf %>% filter(str_detect(demog, "Female")) %>% demogbar()
```
  
  
```{r}
bar_data_hf %>% filter(str_detect(demog, "Male")) %>% demogbar()
```
<br>
<br>


### Black Female Baseline

```{r}
model2_bf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 1) +
               agedummy +
               educ +
               citizen,
             df)
}
```

```{r}
model2_protect_bf <- model2_bf(postprotect)
model2_booker_bf <- model2_bf(postbooker)
model2_gall_bf <- model2_bf(postgall)
model2_report_bf <- model2_bf(postreport)
model2_present_bf <- model2_bf(present)
```

```{r}
protect_bf <- extract_coeffs_demog(model2_protect_bf) %>% mutate(period = "Protect")
booker_bf <- extract_coeffs_demog(model2_booker_bf) %>% mutate(period = "Booker")
gall_bf <- extract_coeffs_demog(model2_gall_bf) %>% mutate(period = "Gall")
report_bf <- extract_coeffs_demog(model2_report_bf) %>% mutate(period = "Post-Report")
present_bf <- extract_coeffs_demog(model2_present_bf) %>% mutate(period = "Present")

bar_data_bf <- bind_rows(protect_bf,
                         booker_bf,
                         gall_bf,
                         report_bf,
                         present_bf) %>% 
  mutate(period = fct_relevel(period, "Protect", "Booker", "Gall", "Post-Report", "Present"))

bar_data_bf %>% filter(str_detect(demog, "Female")) %>% demogbar()
```
  
  
```{r}
bar_data_bf %>% filter(str_detect(demog, "Male")) %>% demogbar()
```
